Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets

### 2020-05-24 ans_to_reviewers/answers-to-reviewers.tex

\begin{answers}
\item{\revcom : }\label{r2c1}

I know very little about the AUVs you discuss, and I suspect many of your readers will be in a similar position. Additional contextual information would be helpful

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[2.1.1] What sort of onboard computing do they have? It may not be something like an Intel 2.1 GHz, 4-thread processor, but a more informative description of something most readers relate to would be helpful. Similar information related to RAM and hard drive space would be helpful. What are the computational limits of these machines?\par
\action: We have added more information about AUVs and their computational capabilities in Section 2.1 (Background on Ocean Processes and Observations) to address this. 
\begin{quote}
The AUVs, see Fig. 9, considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF). The in-water operation time capacity depends on survey speed, payload sensors and navigation. Generally this amounts to 8-24 hrs. operation for smaller crafts, and 24-48 hrs. for larger crafts. AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation. 
\end{quote}
\vspace{1em}

\item[2.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par 
\action: Maximum speed and payload are limiting for coverage. Generally this amounts to 8-24 hrs. operation for smaller crafts, and 24-48 hrs. for larger crafts. Time constraints for operations are typically related to costs of ship time and man-hours during operation. Battery capacity is not a challenge for the AUV described in the paper. 
\vspace{1em}

### 2020-05-24 main_new.tex
Updated with macros

@@

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e2}}
\address{David Ginsbourger\\ Idiap Research Institute\\ Rue Marconi 19, 1920 Martigny\\ Switzerland\\
\printead{e3}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\

\printead{e4}}

@@

The AUVs considered here operate at $1$-$3$ m/s in the upper water column. The in-water operation time capacity depends on survey speed, payload sensors and navigation, see Fig. \ref{fig:lauv}. AUVs have limited computational capacity, and they are usually done on a single-board computer (SBC), like the Raspberry Pi or an multicore GPU NVIDIA Jetson TX1.

Surveys with AUVs are usually limited to observations along fixed transects that are pre-programmed by the human operator. This occurs on a spatial grid, called a waypoint graph. A more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig. \ref{fig:envir1}), making the information gathering \emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}. In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to  modify its path at each waypoint based on what has been measured onboard \citep{py10,Rajan12,Rajan12b}.

@@

%\begin{figure}[h]
%\centering
%\includegraphics[width=0.4\textwidth]{Data-driven.png}
%\label{fig:auv_alone}
%\caption{The concept of data-driven sampling.}
%\end{figure}

%The goal is effectively map the extension of the river plume given this

%, we can formulate this question as follows: the ocean region  defined by the boundary between water bodies, separated by a characteristic temperature gradient that is smaller than $T=5$\textdegree C and a salinity concentration lower than $S=30$ [g/kg]. A goal is then to improve our predictive capabilities to answer this question better.
%For example,  (examples are shown in Fig. \ref{fig:river_fronts}). However,

% In doing so, the aim is to increase the knowledge about the uncertain ocean environment through adaptive sampling

\section{Excursion Sets and Integrated Bernoulli Variance}
\label{sec:ESEP}

\cite{adler2000excursion,adler2009random}.

Our objective is to leverage statistical tools to characterize a river
plume, focusing on spatial separation of cold freshwater from the
river and warmer saline waters of the fjord. ES and EPs are useful
starting points to measure the ability to classify water masses. The
goal of experimental sampling is to improve this
characterization, as quantified via changes in the EPs.

We consider a lateral domain near the sea surface, with locations
$\bx \in \mathcal{M} \subset \mathbf{R^2}$. The two variables of
interest, temperature and salinity, are defined as random processes:
$\xi_T(\bx)$ denotes the temperature (in $^o C$) and $\xi_S(\bx)$
denotes the salinity (in $g/kg$) at location $\bx$. The generalization
to more than two variables is given in Section \ref{gen_k}, and the
adaptive sampling strategies are described in Section
\ref{sec:heuristics}.

We let $t_T$ be a threshold for temperature and $t_S$ a threshold for
salinity. The joint ES is then defined by:
\begin{equation}\label{ES}
     \mbox{ES} = \{\bx \in \mathcal{M} | \xi_T(\bx) \leq t_T,\xi_S(\bx) \leq t_S\}.
\end{equation}
The excursions could also be defined as larger than a threshold, or
within a boundary of limits for the two variables; calculations and
interpretations will be similar in these cases.

The ES holds random indicator variables while the associated EPs are: 
\begin{equation}\label{eq:prob}
  p(\bx) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S), \hspace{3mm} \bx \in \mathcal{M}.
\end{equation}
If the EPs in Eq. (\ref{eq:prob}) are
close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean; it is, however,
challenging if the probability is close to $0.5$. 
The IBV, see \cite{bect2019}, is defined as:
\begin{equation}\label{measVAR}
    V = \int_{\bx \in \mathcal{M}} p(\bx) \left(1-p(\bx)\right) d\bx.
\end{equation}

@@

Our goal is to construct adaptive sampling strategies based on
continuous evaluation of the EPs. As the EPs provide a description of
the boundary uncertainty, the strategies can effectively use this
description to prioritize sampling efforts at locations that are
ambiguous, making exploration more effective. It is natural to
evaluate the expected reduction in the IBV, conditioned on the survey
data. The expected updated IBV is:

@@

@@

\section{Gaussian Processes and Excursion Probabilities}
\label{sec:GP_EP}

@@

\subsection{Bivariate GPs}

\subsection{Data Model and Updating of GPs}

The GP mean and covariance in Eq. (\ref{prior}) are updated in the
model when more information is available. The updated distribution for
temperature and salinity variables on the spatial grid, given survey
data $\by$, is Gaussian with mean and covariance:
\begin{eqnarray}\label{gp_upd}
  \bm &=& \bm(\by) = \bmu+\bSigma \bG^t (\bG \Sigma \bG^t+\bR)^{-1}(\by-\bG \bmu),  \\
  \bA &=& \bSigma - \bSigma \bG^t (\bG \bSigma \bG^t+\bR)^{-1} \bG
          \bSigma.\nonumber
\end{eqnarray}

@@

\section{Uncertainty reduction in excursion sets}
\label{sec:sur}

\subsection{EPs for GPs}

@@

\subsection{Expected IBV for Bivariate Models}

We will next derive closed-form solutions for the expectation in
Eq. (\ref{sur}). The location index $\bx$ is suppressed to avoid
overly complex notation. We present new results for computing the
expectation in the integral of Eq. (\ref{sur}).

A critical first element in the derivation is using the GP model and the
linear dependence on $\by$ in the conditional mean in
Eq. (\ref{gp_upd}). This results in a closed form Gaussian
distribution for the conditional mean in Eq. (\ref{dist_mxi}). The
high-dimensional inner integral in Eq. (\ref{sur}) then reduces to a
bivariate integral \citep{bhattacharjya2013value, chevalier2014fast},
with an expectation over $\bm=(m_{T},m_{S})^t$.  We standardize the two
variables in Eq. (\ref{sur}) as   $Z_T=\frac{\xi_T-m_{T}}{a_{T}}$,
$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$, with
$\mbox{Corr}(Z_T,Z_S)=a_{T,S}/(a_{T} a_{S})=\eta_{T,S}$, and we can
then rephrase the probability as: 
\begin{eqnarray}\label{ytom}
   P(\xi_T \leq t_T, \xi_S \leq t_S|\by) &=& P \left( Z_T \leq\frac{t_T-m_{T}}{a_{T}}, Z_S \leq\frac{t_S-m_{S}}{a_{S}}|\by \right) \nonumber \\
   &=& \Phi_2 \begin{pmatrix} 
\begin{bmatrix} \frac{t_T-m_{T}}{a_{T}}\\
\frac{t_S-m_{S}}{a_{S}}
\end{bmatrix};
 \begin{bmatrix} 0\\
0
\end{bmatrix},\begin{bmatrix}
1 & \eta_{s,t}  \\
\eta_{s,t}   & 1  
\end{bmatrix}\end{pmatrix} \nonumber \\
&=& P(\xi_T \leq t_T, \xi_S \leq t_S|\bm).
\end{eqnarray}
Since the expression only depends on $\by$ via its affine function
$\bm$, the expectation in Eq. (\ref{sur}) is reduced to a bivariate
integral over $\bm$.  Introducing $\pi(\bm)$ for the density function
of $\bm$, the expected Bernoulli variance becomes:
\begin{eqnarray}
\label{eq:var2}
E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \nonumber \\
E_{\bm}(p(1-p)) & = & \int p(1-p) \pi(\bm) d\bm, \nonumber \\
 &=& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm)  \pi(\bm) d\bm \nonumber  \\
&-& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) \pi(\bm) d\bm. 
\end{eqnarray}
We will next extend results of \cite{chevalier2014fast} to solve for
these two parts (see also \cite{stroh}).

@@

For the second part of the expression Eq.(\ref{eq:var2}), standardization
implies that: 
\begin{eqnarray}\label{part1:phi4}
&& \hspace{6mm} \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) p(\bm) d\bm =  \\
&P& \hspace{-3mm} \left( Z_{1,T} \leq \frac{t_T-m_{T}}{a_{T}}, 
Z_{1,S} \leq \frac{t_S-m_{S}}{a_{S}},Z_{2,S} \leq \frac{t_T-m_{T}}{a_{T}}, 
Z_{2,S} \leq \frac{t_S-m_{S}}{a_{S}} \right). \nonumber
\end{eqnarray}
Here, $\bZ_1=(Z_{1,T},Z_{1,S})$, $\bZ_2=(Z_{2,T},Z_{2,S})$ are
independent bivariate zero-mean and unit-variance Gaussian vectors,
both with element-wise correlation $\eta_{T,S}$. Moreover, $\bZ_1$ and
$\bZ_2$ are chosen to be independent of $\bm$, hence:

\begin{equation}
\small
\label{eq:22}
    \left(
    \begin{array}{cccccc}
         Z_{1,T} \\
         Z_{1,S} \\
         Z_{2,T} \\
         Z_{2,S} \\
         m_{T} \\
         m_{S}
    \end{array}
    \right)
\sim N \left(
\left(
    \begin{array}{cccccc}
    0 \\
    0 \\
    0 \\
    0 \\
          \mu_{T} \\
         \mu_{S} 
    \end{array}
    \right),
   \left[ 
      \begin{array}{cccccc}
        1 & \eta_{T,S} & 0 & 0 & 0 & 0  \\
        \eta_{T,S} & 1 & 0 & 0 & 0 & 0 \\
        0 & 0 & 1 & \eta_{T,S} & 0 & 0 \\
        0 & 0 & \eta_{T,S} & 1 & 0 & 0 \\
        0 & 0 & 0 & 0 & \psi^2_{T} & \psi_{T,S} \\
        0 & 0 & 0 & 0 & \psi_{T,S} & \psi^2_{S} 
    \end{array}
\right]
\right).
\vspace{0.2cm}
\end{equation}

The solution to the second part of Eq. \eqref{eq:22} can be found by
evaluating a four-variable cumulative distribution function associated
with Eq. (\ref{part1:phi4}), accounting for the mean, variance and
covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear
combinations. Using vector-matrix formulations, we define the linear
combinations:
\begin{equation}
\label{Vcomb}
    \bV=\left[
    \begin{array}{cccccc}
        a_T & 0 & 0 & 0 & 1 & 0 \\
         0 & a_S & 0 & 0 & 0 & 1 \\
         0 & 0 & a_T & 0 & 1 & 0\\
         0 & 0 & 0 & a_S & 0 & 1\\
    \end{array}
    \right] 
    \left(
    \begin{array}{cccccc}
         Z_{1,T} \\
         Z_{1,S} \\
         Z_{2,T} \\
         Z_{2,S} \\
         m_{T} \\
         m_{S}
    \end{array}
    \right),
\end{equation}
which in turn then becomes:
{\small
\begin{eqnarray}
\label{part1:phi4_2}
&&P \left( Z_{1,T} \leq \frac{t_T-m_{T}}{a_{T}}, 
Z_{1,S} \leq \frac{t_S-m_{S}}{a_{S}}, Z_{2,S} \leq \frac{t_T-m_{T}}{a_{T}}, Z_{2,S} \leq \frac{t_S-m_{S}}{a_{S}} \right)\\
&=&  \Phi_4 
\begin{pmatrix} 
\begin{bmatrix} t_T\\
t_S \\
t_T \\
t_S
\end{bmatrix};
E( \bV),Var(\bV)
\end{pmatrix} \nonumber.
\end{eqnarray}
}
The multivariate cumulative probabilities in two and four dimensions
are computationally efficient using methods developed by
\cite{genz2009computation}.  In the end, the first part in
Eq. (\ref{two_parts0}) and the second in Eq. (\ref{part1:phi4}) are
computed for each location on the grid and summed over
$\bx \in \mathcal{M}_g$, to get the design evaluation criterion in
Eq. (\ref{sur}).

\subsection{Generalization to the Multivariate Case}
\label{gen_k}

In our application domain with the variables temperature and salinity,
the solutions to the complex integral equations for the expected IBV
are reduced to calculating bivariate and four-variate cumulative
distribution functions for Gaussian vectors. For the more general
case, one would be interested in $K$ random fields. In the ocean
sciences, the variables could include chlorophyll or other
bio-geochemical quantities in addition to temperature and salinity.

@@

\subsection{An Example of IBV using Temperature and Salinity}

@@

>>>>

% New Marcos.
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{cleveref}
\input{macros}

@@

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e2}}
\address{David Ginsbourger\\ Idiap Research Institute\\ Rue Marconi 19, 1920 Martigny\\ Switzerland\\
\printead{e3}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
\printead{e4}}

@@

The AUVs, see Fig. \ref{fig:lauv}, considered here operate at $1$-$3$ m/s in the upper water column, where they are free to move in six degrees of freedom (6 DOF). The in-water operation time capacity depends on survey speed, payload sensors and navigation. Generally this amounts to 8-24 hrs. operation for smaller crafts, and 24-48 hrs. for larger crafts. AUVs have limited computational capacity, as they typically use single-board computers (SBCs), like the Raspberry Pi or a multicore GPU NVIDIA Jetson TX1 (quad-core 1.91 GHz 64-bit ARM machine, a 2-MB L2 shared cache, and 4 GB of 1600 MHz DRAM) to do computation.

Surveys with AUVs are usually limited to observations along fixed transects that are pre-scripted in mission plans created manually by a human operator. The mission can be specified operating on a scale of hundreds of meters to tens of kilometers depending on the scientific context. Faced with limited coverage capacity, a more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig. \ref{fig:envir1}), making the information gathering \emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}. This usually occurs on a spatial grid, called a waypoint graph. In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to modify its path at each waypoint based on what has been measured onboard \citep{py10,Rajan12,Rajan12b}.

@@

\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
\label{sec:ESEP}

\subsection{Excursion sets and Integrated Bernoulli variance}
\label{ES}

Our objective is to leverage statistical tools onboard autonomous vehicles to characterize a river
plume, focusing on spatial separation of cold freshwater from the
river and warmer saline waters of the fjord. The characterization of uncertainties on the ES is a useful
goal of this experimental sampling.

In order to formalize the problem in mathematical terms,
let us consider a multivariate (i.e., vector-valued) Gaussian random field $(\gp[\x])_{\x \in D}$, where $D$ is an arbitrary set (typically in geostatistical studies, a subset of interest of the ambient Euclidean space or of a sphere therein) and the dimensionality of the output $\no \geq 1$ is arbitrary. Assume furthermore that the set of interest is characterized in terms of a subset of values of the variables, say $\T \subset \R^{\no}$, and denote the corresponding set of index points by 
$$
\es=\gp^{-1}(\T)=\{\x \in D: \gp[\x] \in \T\}.
$$
Our goal here is to develop approaches to quantify and improve the characterization of uncertainties on such sets in the context of multivariate random fields. 
While some aspects of the developed
approaches do not call for a specific form of $\T$, we will often stick for
simplicity to the case of orthants ($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where $t_1,\dots, t_{\no} \in \R$) as this will allow efficiently calculating several key quantities. Note that changing some $\leq$ inequalities to $\geq$ ones would lead to immediate adaptations. We will keep the abstract notation $\T$ and it shall be clear from the context when specific properties following the orthant assumption come into play. Note that assuming that $\gp$ has continuous trajectories (almost surely) and $T$ to be closed, then $\es$ can be cast as a Random Closed Set \cite{Molchanov2005}. In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.

For our application, we let $t_1$ be a threshold for temperature and $t_2$ a threshold for
salinity. The ES is then defined by $\mbox{ES} = \{\bx \in \mathcal{M} | \xi_T(\bx) \leq t_T,\xi_S(\bx) \leq t_S\}$.
This ES holds random indicator variables while the associated EPs are $p(\bx) = P(\xi_T(\bx) \leq t_T, \xi_S(\bx) \leq t_S)$, $\bx \in \mathcal{M}$.
If the EPs are
close to $1$ or $0$ at a given location, it is easy to classify water mass at this point as either river or ocean. It is, however,
challenging if the probability is close to $0.5$.

Relate to ES and refer to 
\cite{adler2000excursion,adler2009random}.

Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution of $\mes(\es)$ through its moments.

\medskip

%That $\mu(\es)$ is a well-defined random variable follows from
\begin{propo}
	\label{propo1}
	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
	For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for any $r\geq 1$, 
	\begin{equation*}
	\begin{split}
	\mathbb{E}[\mes(\es)^r]
	&=\int_{D^{r}} \jointExcuProb 
	\productMeasure
	,
	\end{split}
	\end{equation*}

	where the product measure is denoted as
	$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$. 
	%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
	Here $\gp$ is defined on $D$, and for
	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
	\medskip

	In the particular case where $\gp$ is a multivariate Gaussian random field 
	%and denoting the mean and covariance matrix of $\gp[\bm{u}]$ by $\meanUU\in\mathbb{R}^{p\times r}$ and $\covUU$, respectively, 
	we have 
	\begin{align*}
	\jointExcuProb = \mathcal{N}_{\no r}(T^r; \meanUU, \covUU),
	\end{align*}
	where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ and covariance matrix $\covUU$, respectively defined blockwise by  
	%where both are defined blockwise by 
	\begin{align*}
	\meanUU&=\begin{pmatrix}\mu(u^{(1)})\\ \vdots\\ \mu(u^{(r)})\end{pmatrix}
	\in \mathbb{R}^{\no r}, \\
	\text{and } \covUU &= \begin{pmatrix}
	\cov(\gp[u^{(1)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(1)}],
	\gp[u^{(r)}])\\
	\vdots & & \vdots\\
	\cov(\gp[u^{(r)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(r)}],
	\gp[u^{(r)}])\\
	\end{pmatrix}\in \mathbb{R}^{pr\times pr},
	\end{align*}
	each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times \no$. 
	%dimensional Gaussian random vector. 
	Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no r$-dimensional Gaussian probability density function 
	$\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as 
	%$\varphi_{\no \times r}$ as
	%$\bm{u}\in D^r$ such that $\gp[\bm{u}]$ be non-degenerate, the integrand can be expressed in terms of the $r \times \no$-dimensional Gaussian distribution, namely 
	%cumulative distribution function, namely 
	\begin{equation*}
	\begin{split}
	&\jointExcuProb
	=
	\int_{T^r} \varphi_{\no  r}\left(\bm{v};
	~\meanUU, \covUU\right) 
	\mathrm{d}\bm{v},
	\end{split}
	\end{equation*}
	In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
	the latter probability directly writes in terms of the multivariate Gaussian
	cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
	this time by the way without requiring $\covUU$ %the latter 
	to be non-singular: 
	\begin{equation*}
	\begin{split}
	\jointExcuProb 
	&=
	\varPhi_{\no r}\left(\bm{t};~\meanUU, \covUU\right),
	\end{split}
	\end{equation*}
	where we have used the notations 
	$t=(t_1,\dots,t_{\no}%,..., t_1, ..., t_p
	)\in\mathbb{R}^{\no}$, $1_{r}=(1,\dots,1)\in \R^{r}$, and 
	$\bm{t}=1_{r}\otimes \bm{t}=(t_1,\dots,t_{\no},\dots,t_1,\dots,t_{\no}) 
	\in \R^{\no r}$.
\end{propo}

The multivariate cumulative probabilities in two and four dimensions
are computationally efficient using methods developed by
\cite{genz2009computation}.

Proposition~\ref{propo1} enables in turn calculating centred moments of $\mes(\es)$. 
The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is probably the most relevant of them for the considered purposes. We have
\begin{equation*}
\begin{split}
\emv
&=\int_{D^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right) 
d\mes^{\otimes}(u, v)\\
&-\left( \int_{D} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*} 
which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
\dots \times (-\infty, t_{\no}]$ to 
\begin{equation*}
\begin{split}
\emv
%\operatorname{Var}[\mes(\es)]
&=\int_{D^2} 
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)), 
K((u,v),(u,v))
\right) 
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
&-\left( \int_{D} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
%
Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $D^2$. In contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves solely and integral on $D$ and can be expanded in our present settings as follows
\begin{equation*}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{D}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{D}
\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) 
-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u). 
\end{split}
\end{equation*}

\begin{propo}
	\label{propo2}
	Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
	and $\covN, \covV$ be two covariance matrices in
	$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
	Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
	\begin{equation*}
	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; \covN \right)^h \right]
	=
	\varPhi_{ph}
	\left(
	\bm{a}
	;~
	\bm{\Sigma}
	\right),
	\end{equation*}
	where the vector $\bm{a} \in \R^{p h}$ is defined as 
	$\bm{a} := 1_h\otimes a = 
	\left(a, \dots , a
	\right)'$
	and the $p h\times p h$ covariance matrix is given by
	$\bm{\Sigma} := 
	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
\end{propo}

	In blockwise representation, $\bm{\Sigma}$ can be expressed as follows: 
	\begin{align*}
	% \bm{\Sigma}&=
	\begin{pmatrix}
	\covN & &\\
	& \ddots &\\
	&   & \covN
	\end{pmatrix}
	+
	\begin{pmatrix}
	B\covV B' & \dots & B\covV  B'\\
	\vdots & & \vdots\\
	B\covV B' & \dots & B\covV B'\\
	\end{pmatrix}
	\end{align*}

In the Appendix we also provide a more general derivation which holds for more general criterian than IBV, in partical for volumes of sets, meaning correlations between many locations meaning sums of indicator variables in Prop 1.

\subsection{Sampling Designs and Expected Integrated Bernoulli Variance}

Next we investigate how sampling designs are chosen to minimize the expected Integrated Bernoulli variance (EIBV).

In order to be able to calculate them in semi-analytical forms relying on integrals of multivariate Gaussian cumulative distribution functions, we will appeal to some key intermediate results presented next.

(MODIFY NOTATION HERE - BUT NEED THIS $y$ IN HERE)

@@

Our goal is to construct adaptive sampling strategies based on
continuous evaluation of the EPs. As the EPs provide a description of
the boundary uncertainty, the strategies can effectively use this
description to prioritize sampling efforts at locations that are
ambiguous, making exploration more effective. It is natural to
evaluate the expected reduction in the IBV, conditioned on the survey
data. The EIBV is:

(MODIFY NOTATION HERE)

@@

REUSE PROPOSITION RESULTS

@@

@@

(MODIFY NOTATION HERE)

The GP mean and covariance in Eq. (\ref{prior}) are updated in the
model when more information is available. The updated distribution for
temperature and salinity variables on the spatial grid, given survey
data $\by$, is Gaussian with mean and covariance:
%\begin{eqnarray}\label{gp_upd}
%  \bm &=& \bm(\by) = \bmu+\bSigma \bG^t (\bG \Sigma \bG^t+\bR)^{-1}(\by-\bG \bmu),  %\\
%  \bA &=& \bSigma - \bSigma \bG^t (\bG \bSigma \bG^t+\bR)^{-1} \bG
%          \bSigma.\nonumber
%\end{eqnarray}

@@

@@

MODIFY AND REUSE PROPOSITIONS

A critical first element in the derivation is using the GP model and the
linear dependence on $\by$ in the conditional mean in
Eq. (\ref{gp_upd}). This results in a closed form Gaussian
distribution for the conditional mean in Eq. (\ref{dist_mxi}). The
high-dimensional inner integral in Eq. (\ref{sur}) then reduces to a
bivariate integral \citep{bhattacharjya2013value, chevalier2014fast},
with an expectation over $\bm=(m_{T},m_{S})^t$.  We standardize the two
variables in Eq. (\ref{sur}) as   $Z_T=\frac{\xi_T-m_{T}}{a_{T}}$,
$Z_S=\frac{\xi_S-m_{S}}{a_{S}}$, with
$\mbox{Corr}(Z_T,Z_S)=a_{T,S}/(a_{T} a_{S})=\eta_{T,S}$, and we can
then rephrase the probability as: 
\begin{eqnarray}\label{ytom}
   P(\xi_T \leq t_T, \xi_S \leq t_S|\by) &=& P \left( Z_T \leq\frac{t_T-m_{T}}{a_{T}}, Z_S \leq\frac{t_S-m_{S}}{a_{S}}|\by \right) \nonumber \\
   &=& \Phi_2 \begin{pmatrix} 
\begin{bmatrix} \frac{t_T-m_{T}}{a_{T}}\\
\frac{t_S-m_{S}}{a_{S}}
\end{bmatrix};
 \begin{bmatrix} 0\\
0
\end{bmatrix},\begin{bmatrix}
1 & \eta_{s,t}  \\
\eta_{s,t}   & 1  
\end{bmatrix}\end{pmatrix} \nonumber \\
&=& P(\xi_T \leq t_T, \xi_S \leq t_S|\bm).
\end{eqnarray}
Since the expression only depends on $\by$ via its affine function
$\bm$, the expectation in Eq. (\ref{sur}) is reduced to a bivariate
integral over $\bm$.  Introducing $\pi(\bm)$ for the density function
of $\bm$, the expected Bernoulli variance becomes:
%\begin{eqnarray}
%\label{eq:var2}
%E_{\by}(p(1-p)) &=& E_{\bm}(p(1-p)), \hspace{3mm} p=P(\xi_T \leq t_T, \xi_S \leq %t_S|\bm) \nonumber \\
%E_{\bm}(p(1-p)) & = & \int p(1-p) \pi(\bm) d\bm, \nonumber \\
% &=& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm)  \pi(\bm) d\bm \nonumber  \\
%&-& \int P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) P(\xi_T \leq t_T, \xi_S \leq t_S|\bm) %\pi(\bm) d\bm. 
%\end{eqnarray}
We will next extend results of \cite{chevalier2014fast} to solve for
these two parts (see also \cite{stroh}).

@@

@@

\subsection{An Example of EIBV using Temperature and Salinity}

@@

\section*{Appendix}

Proof of Proposition 1:

\begin{proof}
	That $\mes(\es)$ defines indeed a random variable follows from Fubini's theorem 
	relying on the joint measurability of 
	$(\x, \omega) \to \mathbbm{1}_{\es(\omega)}(\x)$, 
	%= \mathbf{1}_{\xi_{\x}(\omega) \in T}$ and Fubini's theorem.  
	itself inherited from the assumed measurability for  
	$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robins1944}, we find that 
	\begin{equation*}
	\begin{split}
	\mathbb{E}[\mes(\es)^{r}]
	&=\mathbb{E}\left[\left(\int_{D} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
	=\mathbb{E}\left[ \prod_{i=1}^{r} \left(
	\int_{D} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)}) 
	\right) \right] \\
	&
	=
	%\mathbb{E}\left[
	%\int_{D^{r}}
	%\prod_{i=1}^{r}
	%\mathbf{1}_{\gp[\uu^{(i)}] \in T}
	%\productMeasure
	%\right]
	%=
	\mathbb{E}\left[
	\int_{D^{r}}
	\mathbbm{1}_{\gp[\uu^{(1)}] \in T,\dots, \gp[\uu^{(r)}]  \in T}
	\productMeasure
	\right]
	%\\
	%&=\int_{D^{r}} \mathbb{E}\left[\mathbf{1}_{\gp[\bm{u}] \in T^r} \right] \productMeasure \\
	%&
	=\int_{D^{r}}
	\jointExcuProb 
	\productMeasure.
	\end{split}
	\end{equation*}   
	The rest consists in expliciting the probability of $T\times \dots \times T$ under the multivariate Gaussian distribution of  
	$\left(\gp[u^{(1)}], \dots,  \gp[u^{(r)}] \right)$. 
\end{proof}

Proof of Proposition 2:
